### 2013-01-25 patchy-selection-paper.tex
/* Patchy Selection paper */

\subsubsection{Time to the establishment of a locally adaptive allele
  due to migrational influx}
\label{ss:patchymigration}
Now suppose that there are two patches, each of area $A$, of the poor
habitat type separated by distance $R$. Consider the case where the
$B$ allele has arisen and become established in the first patch, but has not yet appeared in the second.
As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
has probability approximately $p(x)$ of fixing locally in this patch,
whether the allele is a new mutation or is descended from the occupants of the first patch.
Due to mutation--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
present in the regions where they are disadvantageous;
the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
although the equations are again not particularly helpful.
However, we do know that if $q(r)$ is the expected proportion present at \gc{migration}--selection balance at distance $r$ from the first patch
in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
where $C$ is a constant depending on the structure of the populations
and the selection coefficients. \gc{comment on size of C? Should we
  also use this approx. to get the approx. rate of establishing mutations
  arising outside of the patch, above. Feel like in the mutation rate
  part we could pull $s_B$ out front of the rate as well.}

@@

Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
supposing that patches are large enough for local fixation to be stable. If the alleles were not advantageous in the second patch,
the average number of mutant alleles present in a small area $\Delta$
near the second patch, due
to migration from the first patch, would be $\approx \rho E \exp( -
|\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
sufficiently large to give an allele a $2s_B$ probability of becoming established, the rate at which migrants find
themselves in the patch and become established  is approximately
\begin{equation}
\lambda_{Mig,R} = \rho E s_B \exp( -|\sqrt{s_m} R/\sigma| ).
\end{equation}
\gc{Is this $A^2$ right here?. Do we want to use $q(R)$ here?}
If this establishment rate is low, the time it takes
a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$
Assuming that the time it takes for a allele from the first patch to
become established in the second patch is short,
then  the expected waiting time until some migrant allele fixes in the
second patch will be $T_{Mig} = 1/\lamda_{Mig,R} $.

>>>>

\subsubsection{Time to the establishment of a locally adaptive allele
  due to migrational influx}
\label{ss:patchymigration}
Now suppose that there are two patches, each of area $A$, of the poor
habitat type separated by distance $R$. Consider the case where the
$B$ allele has arisen and become established in the first patch, but has not yet appeared in the second.
As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
has probability approximately $p(x)$ of fixing locally in this patch,
whether the allele is a new mutation or is descended from the
occupants of the first patch.

For the moment we will ignore the
possibility of mutation generating a $B$ allele can concentrate on the
flow of migrants from patch $1$ to patch $2$.
Due to \gc{migration}--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
present in the regions where they are disadvantageous;
the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
although the equations are again not particularly helpful.
However, we do know that if $q(r)$ is the expected proportion present at \gc{migration}--selection balance at distance $r$ from the first patch
in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
where $C$ is a constant depending on the structure of the populations
and the selection coefficients. \gc{comment on size of C? Should we
  also use this approx. to get the approx. rate of establishing mutations
  arising outside of the patch, above. Feel like in the mutation rate
  part we could pull $s_B$ out front of the rate as well.}

@@

Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
supposing that patches are large enough for local fixation to be stable. If the alleles were not advantageous in the second patch,
the average number of mutant alleles present in a small area $\Delta$
near the second patch, due
to migration from the first patch, would be $\approx \rho E \exp( -
|\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
sufficiently large to give an allele a $2s_B$ probability of becoming established, the rate at which migrants find
themselves in the patch and become established  is approximately
\begin{equation}
\lambda_{Mig,R} = \rho E s_B \exp( -|\sqrt{s_m} R/\sigma| ).
\end{equation}
\gc{Is this $A^2$ right here?. Do we want to use $q(R)$ here?}
If this establishment rate is low, the time it takes
a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$
Assuming that the time it takes for a allele from the first patch to
become established in the second patch is short,
then  the expected waiting time until some migrant allele fixes in the
second patch will be $T_{Mig} = 1/ \lambda_{Mig,R} $.
